Temporality of the Victorian Hospital Patient Transfer Network

### 2025-06-30 reports/paper.qmd
add monday night

We use admissions data sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official]. This contains line-listed admissions for patients in Victoria, Australia, covering 338 healthcare facilities that vary in size, case mix and services.
the dataset contains 26,796,407 individual admissions across 5,471893 unique patients, from 1 January 2011 through to 6 December 2019 (3631 days).
Because there is no unique healthcare identifier in Victoria, Australia, linkage is performed on the VAED dataset by the Centre for Victorian Data Linkage, in order to build information on succesive admissions of patients. As a result of this linkage and hospital administrative procedures, there can be overlapping admissions, where a patient admits at a facility before being discharged at a prior facility. These overlaps are removed by constructing new admissions that are of minimal non-overlapping duration.
Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework.

@@

To motivate the following methods, we first characterise temporal trends in the data.

Firstly, we examine the duration of stays

@@

In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. Qualitatively, there are mechanisms implicit in the data that are not previously well-modelled, and may be of importance to constructing robust control measures. Critically, we notice that the distribution of readmission times (the time from a discharge of a given patient from a facility to next admission of that patient at any facility), as plotted in @fig-readmission-time, are not well modelled by an exponential distribution, i.e. poorly approximated by Markov processes.

@@

>>>>

We use admissions data sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official]. This contains line-listed admissions for patients in Victoria, Australia, covering 338 healthcare facilities that vary in size, case mix and services.
the dataset contains 26,796,407 individual admissions across 5,471893 unique patients, from 1 January 2011 through to 6 December 2019 (3631 days).
Because there is no unique healthcare identifier in Victoria, Australia, linkage is performed on the VAED dataset by the Centre for Victorian Data Linkage, in order to build information on succesive admissions of patients. As a result of this linkage and hospital administrative procedures, there can be overlapping admissions, where a patient admits at a facility before being discharged at a prior facility. These overlaps are removed by constructing new admissions that are of minimal non-overlapping duration.
Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework, and it characterises the structure of the healthcare system.

@@

The temporal characterisation of the Victorian healthcare system is less well-explored.
For example, the amount of patient movement in the system fluctuates temporally, as shown in @fig-movement-over-time. 
There is an increasing amount of transfers over time, and there are dips in transfers related to public holidays in the state of Victoria. This implies that we should have at least an inhomogenous Poisson process governing the rate of transfers between facilites over time.

The underlying admission duration is ___ distributed (see @fig-stay-duration), and the time between admissions (i.e. the transfer period) is power-law distributed (see @fig-readmission-time). Thus, a fully Markovian POisson process on a static network is unlikely to be a good representation of the movement in the network.

@@

@@

## Network Models
